---
title: "Cellphone Customer Churn Prediction"
output: html_notebook
---


Loading the Dataset
```{r}
library(readxl)
Cellphone <- read_excel("D:/GL_PGP/Assignments/Logistic Regression/Cellphone.xlsx", 
    sheet = "Data")
View(Cellphone)
```


Libraries needed
```{r}
library(readxl)
library(psych)
library(dplyr)
library(tidyverse)
library(DataExplorer)
library(ggplot2)
library(esquisse)
library(corrplot)
library(blorr)
library(car)
library(caret)
library(ROCR)

```



Basic Info
```{r}
head(Cellphone)
tail(Cellphone)
str(Cellphone)
summary(Cellphone)
```


```{r}
library(psych)
library(dplyr)
describe(Cellphone)
glimpse(Cellphone)
```


Check for NA values
```{r}

anyNA(Cellphone)
sum(is.na(Cellphone))
colSums(is.na(Cellphone))
```


```{r}
library(tidyverse)
Cellphone %>% map(~ sum(is.na(.)))
```


```{r}
Cellphone_num = Cellphone
```


Converting 'Churn', 'Contract Renewal' and 'Data Plan' to factors and adding levels
```{r}

Cellphone$Churn = as.factor(Cellphone$Churn)
levels(Cellphone$Churn) = c("No","Yes")
Cellphone$ContractRenewal = as.factor(Cellphone$ContractRenewal)
levels(Cellphone$ContractRenewal) = c("No", "Yes")
Cellphone$DataPlan = as.factor(Cellphone$DataPlan)
levels(Cellphone$DataPlan) = c("No", "Yes")


summary(Cellphone)

```


```{r}
library(DataExplorer)
library(ggplot2)
library(esquisse)

create_report(Cellphone)
```


Exploratory Data Analysis
Churn

```{r}
prop.table(table(Cellphone$Churn))
```

```{r}
#esquisser()

ggplot(Cellphone) +
 aes(x = Churn) +
 geom_bar(fill = "#ed7953") +
 theme_minimal()
```


```{r}

plot_boxplot(Cellphone, by = "Churn", 
             geom_boxplot_args = list("outlier.color" = "red"))

```



Accounts Week
```{r}
ggplot(Cellphone) +
 aes(x = AccountWeeks, colour = Churn) +
 geom_histogram(bins = 30L, fill = "#0c4c8a") +
 scale_color_hue() +
 theme_minimal()
```

```{r}
ActWeeks <- cut(Cellphone$AccountWeeks, breaks = seq(0, 250, by = 25))
ggplot(Cellphone, aes(ActWeeks, fill = Churn)) + geom_bar(position="dodge")
```


```{r}
ggplot(Cellphone) +
 aes(y = AccountWeeks, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```



```{r}
ggplot(Cellphone) +
 aes(y = AccountWeeks) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()

```


Contract Renewal
```{r}
ggplot(Cellphone, aes(ContractRenewal, fill = Churn)) + geom_bar(position="dodge")
```




Data Plan
```{r}
ggplot(Cellphone, aes(DataPlan, fill = Churn)) + geom_bar(position="dodge")
```



Data Usage
```{r}
ggplot(Cellphone) +
 aes(y = DataUsage) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```



```{r}
#esquisser()
ggplot(Cellphone) +
 aes(y = DataUsage, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```



Customer Service Callls
```{r}
ggplot(Cellphone) +
 aes(y = CustServCalls) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```



```{r}
ggplot(Cellphone) +
 aes(y = CustServCalls, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```



Day Minutes
```{r}
ggplot(Cellphone) +
 aes(y = DayMins) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```

```{r}
ggplot(Cellphone) +
 aes(y = DayMins, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```



Day Calls
```{r}
ggplot(Cellphone) +
 aes(y = DayCalls) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```



```{r}
ggplot(Cellphone) +
 aes(y = DayCalls, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```



Monthly Charge
```{r}
ggplot(Cellphone) +
 aes(y = MonthlyCharge) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```


```{r}
ggplot(Cellphone) +
 aes(y = MonthlyCharge, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```


Overage Fee
```{r}
ggplot(Cellphone) +
 aes(y = OverageFee) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```



```{r}
ggplot(Cellphone) +
 aes(y = OverageFee, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```


Roaming Minutes
```{r}
ggplot(Cellphone) +
 aes(y = RoamMins) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()
```


```{r}
ggplot(Cellphone) +
 aes(y = RoamMins, fill = Churn) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()
```



Correlation Check
```{r}
library(corrplot)
corrplot(cor(Cellphone_num))
corrplot(cor(Cellphone_num), type = "lower", method = "number")
```




Splitting the data for Train & Test

```{r}
numrow=nrow(Cellphone)
set.seed(1025)

split=sample(seq_len(numrow),ceiling(.7*numrow),replace=F)
train = Cellphone[split,]
test = Cellphone[-split,]

dim(train)
dim(test)
```



```{r}
Model1 <- glm(Churn ~., data=train, family = binomial)
summary(Model1)
```

```{r}
library(blorr)

Model2 = blorr::blr_step_aic_both(Model1)
Model2$model

```


```{r}
Model3 <- glm(Churn ~ CustServCalls + ContractRenewal+ DataPlan + MonthlyCharge + DataUsage + RoamMins, data=train, family = binomial)
summary(Model3)
```

```{r}
library(car)
vif(Model1)
```



```{r}
vif(Model3)
```


```{r}
library(caret)
pred_1 <- predict(Model1, newdata = train, type = 'response')
table(train$Churn, pred_1>0.5)

glm.pred_1 = rep("No", length(pred_1))
glm.pred_1[pred_1 > 0.5] = "Yes"
glm.pred_1 <- as.factor(glm.pred_1)

confusionMatrix(glm.pred_1, train$Churn, positive = "Yes")
```


```{r}

pred_2 <- predict(Model3, newdata = train, type = 'response')
table(train$Churn, pred_2>0.5)

glm.pred_2 = rep("No", length(pred_2))
glm.pred_2[pred_2 > 0.5] = "Yes"
glm.pred_2 <- as.factor(glm.pred_2)

confusionMatrix(glm.pred_2, train$Churn, positive = "Yes")
```

```{r}
prop.table(table(train$Churn))
```



Since there is a huge class imbalance, the sensitivity of the model is very less. Let is try some sampling techniques to improve the sensitivity

```{r}
control1 = trainControl(method = "repeatedcv", number = 5, repeats = 3, sampling = "up", set.seed(1200))
Model4 = train(Churn ~ ., data = train, method = 'glm', trControl = control1, set.seed(2111))
Model4 


pred_3 = predict(Model4, newdata = train)
table(train$Churn, pred_3)



confusionMatrix(pred_3, train$Churn, positive = "Yes")
```


```{r}
control2 = trainControl(method = "repeatedcv", number = 5, repeats = 3, sampling = "smote", set.seed(1200))
Model5 = train(Churn ~ ., data = train, method = 'glm', trControl = control2, set.seed(1502))
Model5 


pred_4 = predict(Model5, newdata = train)
table(train$Churn, pred_4)



confusionMatrix(pred_4, train$Churn, positive = "Yes")
```



```{r}
control3 = trainControl(method = "repeatedcv", number = 5, repeats = 3, sampling = "up", set.seed(1200))
Model6 = train(Churn ~ CustServCalls + ContractRenewal+ DataPlan + MonthlyCharge + DataUsage + RoamMins, data = train, method = 'glm', trControl = control3, set.seed(1502))
Model6 


pred_5 = predict(Model6, newdata = train)
table(train$Churn, pred_5)



confusionMatrix(pred_5, train$Churn, positive = "Yes")
```


```{r}
control4 = trainControl(method = "repeatedcv", number = 5, repeats = 3, sampling = "smote", set.seed(1200))
Model7 = train(Churn ~ CustServCalls + ContractRenewal+ DataPlan + MonthlyCharge + DataUsage + RoamMins, data = train, method = 'glm', trControl = control4, set.seed(1502))
Model7 


pred_6 = predict(Model7, newdata = train)
table(train$Churn, pred_6)



confusionMatrix(pred_6, train$Churn, positive = "Yes")
```



Model Performance - Confusion Matrix on Train Data
```{r}
pred_train_1 = predict(Model4, newdata = train)
table(as.factor(train$Churn),pred_train_1)

pred_train1=with(train,table(as.factor(train$Churn),pred_train_1))
pred_train1

TN_train1 = pred_train1[1,1]
TP_train1 = pred_train1[2,2]
FN_train1 = pred_train1[2,1]
FP_train1 = pred_train1[1,2]

train_acc1 = (TN_train1+TP_train1)/(TN_train1+TP_train1+FN_train1+FP_train1)
train_acc1

train_sens1 = TP_train1/(TP_train1+FN_train1)
train_sens1


train_spec1 = TN_train1/(TN_train1+FP_train1)
train_spec1
```

```{r}
pred_train_2 = predict(Model5, newdata = train)
table(as.factor(train$Churn),pred_train_2)


pred_train2=with(train,table(as.factor(train$Churn),pred_train_2))
pred_train2

TN_train2 = pred_train2[1,1]
TP_train2 = pred_train2[2,2]
FN_train2 = pred_train2[2,1]
FP_train2 = pred_train2[1,2]

train_acc2 = (TN_train2+TP_train2)/(TN_train2+TP_train2+FN_train2+FP_train2)
train_acc2

train_sens2 = TP_train2/(TP_train2+FN_train2)
train_sens2


train_spec2 = TN_train2/(TN_train2+FP_train2)
train_spec2
```

```{r}
pred_train_3 = predict(Model6, newdata = train)
table(as.factor(train$Churn),pred_train_3)

pred_train3=with(train,table(as.factor(train$Churn),pred_train_3))
pred_train3

TN_train3 = pred_train3[1,1]
TP_train3 = pred_train3[2,2]
FN_train3 = pred_train3[2,1]
FP_train3 = pred_train3[1,2]

train_acc3 = (TN_train3+TP_train3)/(TN_train3+TP_train3+FN_train3+FP_train3)
train_acc3

train_sens3 = TP_train3/(TP_train3+FN_train3)
train_sens3


train_spec3 = TN_train3/(TN_train3+FP_train3)
train_spec3
```

```{r}
pred_train_4 = predict(Model7, newdata = train)
table(as.factor(train$Churn),pred_train_4)

pred_train4=with(train,table(as.factor(train$Churn),pred_train_4))
pred_train4

TN_train4 = pred_train4[1,1]
TP_train4 = pred_train4[2,2]
FN_train4 = pred_train4[2,1]
FP_train4 = pred_train4[1,2]

train_acc4 = (TN_train4+TP_train4)/(TN_train4+TP_train4+FN_train4+FP_train4)
train_acc4

train_sens4 = TP_train4/(TP_train4+FN_train4)
train_sens4


train_spec4 = TN_train4/(TN_train4+FP_train4)
train_spec4
```





Prediction On Test Data
Model Performance - Confusion Matrix on Test Data

```{r}
pred_test_1 = predict(Model4, newdata = test)
table(as.factor(test$Churn),pred_test_1)

pred_test1=with(test,table(as.factor(test$Churn),pred_test_1))
pred_test1

TN_test1 = pred_test1[1,1]
TP_test1 = pred_test1[2,2]
FN_test1 = pred_test1[2,1]
FP_test1 = pred_test1[1,2]

test_acc1 = (TN_test1+TP_test1)/(TN_test1+TP_test1+FN_test1+FP_test1)
test_acc1

test_sens1 = TP_test1/(TP_test1+FN_test1)
test_sens1


test_spec1 = TN_test1/(TN_test1+FP_test1)
test_spec1
```


```{r}
pred_test_2 = predict(Model5, newdata = test)
table(as.factor(test$Churn),pred_test_2)


pred_test2=with(test,table(as.factor(test$Churn),pred_test_2))
pred_test2

TN_test2 = pred_test2[1,1]
TP_test2 = pred_test2[2,2]
FN_test2 = pred_test2[2,1]
FP_test2 = pred_test2[1,2]

test_acc2 = (TN_test2+TP_test2)/(TN_test2+TP_test2+FN_test2+FP_test2)
test_acc2

test_sens2 = TP_test2/(TP_test2+FN_test2)
test_sens2


test_spec2 = TN_test2/(TN_test2+FP_test2)
test_spec2
```


```{r}
pred_test_3 = predict(Model6, newdata = test)
table(as.factor(test$Churn),pred_test_3)

pred_test3=with(test,table(as.factor(test$Churn),pred_test_3))
pred_test3

TN_test3 = pred_test3[1,1]
TP_test3 = pred_test3[2,2]
FN_test3 = pred_test3[2,1]
FP_test3 = pred_test3[1,2]

test_acc3 = (TN_test3+TP_test3)/(TN_test3+TP_test3+FN_test3+FP_test3)
test_acc3

test_sens3 = TP_test3/(TP_test3+FN_test3)
test_sens3


test_spec3 = TN_test3/(TN_test3+FP_test3)
test_spec3
```


```{r}
pred_test_4 = predict(Model7, newdata = test)
table(as.factor(test$Churn),pred_test_4)

pred_test4=with(test,table(as.factor(test$Churn),pred_test_4))
pred_test4

TN_test4 = pred_test4[1,1]
TP_test4 = pred_test4[2,2]
FN_test4 = pred_test4[2,1]
FP_test4 = pred_test4[1,2]

test_acc4 = (TN_test4+TP_test4)/(TN_test4+TP_test4+FN_test4+FP_test4)
test_acc4

test_sens4 = TP_test4/(TP_test4+FN_test4)
test_sens4


test_spec4 = TN_test4/(TN_test4+FP_test4)
test_spec4
```





Model Performance- ROC and AUC 
For Model 4- Up-sampling Full Model
```{r}
library(pROC)

test$predict.score=predict(Model4,test)
test$predict.score=predict(Model4,test, type = "prob")
test$predict.score
roc_obj = roc(test$Churn, test$predict.score[,2])


plot(roc_obj, print.auc = T)


```

Remove predicted score before running other models
```{r}

test$predict.score = NULL
```


For Model 5- Smote-sampling Full Model
```{r}
test$predict.score=predict(Model5,test)
test$predict.score=predict(Model5,test, type = "prob")
test$predict.score
roc_obj = roc(test$Churn, test$predict.score[,2])


plot(roc_obj, print.auc = T)
```

Remove predicted score before running other models
```{r}

test$predict.score = NULL
```


For Model 6- Up-sampling Significant Variables Model
```{r}
test$predict.score=predict(Model6,test)
test$predict.score=predict(Model6,test, type = "prob")
test$predict.score
roc_obj = roc(test$Churn, test$predict.score[,2])


plot(roc_obj, print.auc = T)
```

Remove predicted score before running other models
```{r}

test$predict.score = NULL
```


For Model 7- Smote-sampling Significant Variables Model
```{r}
test$predict.score=predict(Model7,test)
test$predict.score=predict(Model7,test, type = "prob")
test$predict.score
roc_obj = roc(test$Churn, test$predict.score[,2])


plot(roc_obj, print.auc = T)
```





```{r}
fram_data_frame1 = data.frame(train_acc1, train_sens1, train_spec1)
names(fram_data_frame1) = c("ACC", "SENS", "SPEC")
fram_data_frame2 = data.frame(train_acc2, train_sens2, train_spec2)
names(fram_data_frame2) = c("ACC", "SENS", "SPEC")
fram_data_frame3 = data.frame(train_acc3, train_sens3, train_spec3)
names(fram_data_frame3) = c("ACC", "SENS", "SPEC")
fram_data_frame4 = data.frame(train_acc4, train_sens4, train_spec4)
names(fram_data_frame4) = c("ACC", "SENS", "SPEC")
fram_data_frame5 = data.frame(test_acc1, test_sens1, test_spec1)
names(fram_data_frame5) = c("ACC", "SENS", "SPEC")
fram_data_frame6 = data.frame(test_acc2, test_sens2, test_spec2)
names(fram_data_frame6) = c("ACC", "SENS", "SPEC")
fram_data_frame7 = data.frame(test_acc3, test_sens3, test_spec3)
names(fram_data_frame7) = c("ACC", "SENS", "SPEC")
fram_data_frame8 = data.frame(test_acc4, test_sens4, test_spec4)
names(fram_data_frame8) = c("ACC", "SENS", "SPEC")

fram_fin =rbind(fram_data_frame1, fram_data_frame2,fram_data_frame3,fram_data_frame4,fram_data_frame5, fram_data_frame6,fram_data_frame7,fram_data_frame8)
row.names(fram_fin) = c('GLM_Full_upsample_train','GLM_Full_smotesample_train','GLM_Sig_upsample_train','GLM_Sig_smotesample_train','GLM_Full_upsample_test','GLM_Full_smotesample_test','GLM_Sig_upsample_test','GLM_Sig_smotesample_test')
fram_fin
```

